---
title: "Files-Folders-Homework"
author: "Patty Vail"
date: "2024-02-09"
output: html_document
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library("fs")

```

## Assignment for Module 3: Files & Folders

To recieve full credit, complete all sections. Upload your knitted Rmd in html, not Rmd, to Canvas. If you are having issues knitting your Rmd, email Shady (shady.kuster@colostate.edu), email David (david.king@colostate.edu), or attend coding club this Friday for help. I'm (shady) happy to meet in person, too, if that is more helpful, but I am likely unable to attend coding club due to prior commitments.  

This assignment is due *February 15th at 11:59 PM*. 


## Conceptual Questions

1) What is the difference between an absolute path and a relative path?

The absolute path gives the full address or location of where your file is. The relative path gives the location of the file that is in relation to where you are CURRENTLY working. 

2) When might you use a relative or an absolute path?

You will use the absolute path when you are setting your working directory. You will use the relative path when you want to know where your current working directory is. 

3) If you get the error "cannot open file 'example.csv': No such file or directory", what things should you check first?

What is you working directory? Where did you put your .csv file? Also, where is the current assignment( IE: Homework Assigment.Rmd) located. You want to make sure you are telling your computer the correct path to access these files. 

## Applied Questions

4) Create the following directory system. Use dir_tree() to output the directory structure of the directory you made. 

Within your cwd (change to a new, empty folder):
ExFolder is empty
Plant-Genera contains two directories: 1) Silene and 2) Arabidopsis
Projects contains a directory called Linkage and two text files: 1) to-do.txt and 2) project-ideas.txt

```{r}
#Question 4
setwd("/Users/pvail/Documents/Courses/Comp Bio/Files and folderes class")
dir_tree("HW Assignment")

```

5) Make sure you have downloaded the HomeworkAssignmentInput.txt file from Canvas. 
a) without modifying the code below and keeping your cwd the same as before (do not modify it), successfully run the following code chunk:

```{r}
getwd()
#Question 5a - do not edit any of this code
paste(readLines("HomeworkAssignmentInput.txt"))

dir_tree(".")

```

b) copy the file to the Silene directory, and read in the file from that directory, keeping your cwd the same as the previous question. 

```{r}
getwd()
#Question 5b 
paste(readLines("Plant-Genera/Silene/HomeworkAssignmentInput.txt")) #fill in path here

dir_tree(".") #keep this here

```

c) now, change your working directory to be able to read the file from the Silene directory.

```{r}
#Question 5c
#add what you need to here
setwd("/Users/pvail/Documents/Courses/Comp Bio/Files and folderes class/HW Assignment/Plant-Genera/Silene")
paste(readLines("HomeworkAssignmentInput.txt")) #do not change this line

dir_tree(".") #keep this here
```

