---
title: "Homework Tidy Verse"
author: "Patricia Vail"
output: html_document
date: "2024-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries

```{r}
library(tidyverse)
```

#### Read in data 

```{r}
getwd()
#Load data
surveys <- read_csv("/Users/pvail/Documents/Courses/Comp Bio/TidyVerse/Class Repo/CM515-course-2024/modules/04_Tidyverse/diversity.csv")
```

### Tidyverse Day 1

#### Q1: What is the class of the object surveys? 

```{r}
#Q1
class(surveys)
str(surveys)
```
Surveys is a data frame 

#### Q2: How many rows and how many columns are in the object?

Rows: 34786
Columns: 13

#### Q2: Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.

```{r}
#Q2
surveys[200,]
```

#### Q3: Create a data.fram (surveys_last) containing only the data in row 200 of the surveys dataset using nrow() instead of the row number. ?nrow

```{r}
#Q3
nrows <- nrow(surveys)
surveys_last <- surveys[nrows,]
```

#### Q4: Change the columns taxa and genus in the surveys data frame into a factor.

```{r}
#Q4
surveys$taxa <- factor(surveys$taxa)
str(surveys)
surveys$genus <- factor(surveys$genus)
str(surveys)
```

#### Q5: Using the functions you learned before

##### Q5 A: How many rabbits were observed? 

75 rabbits observed
```{r}
#Q5A
table(surveys$taxa)
```


##### Q5 B: How many different genera are in the genus column?
There are 26 genera in the genus column
```{r}
#Q5B
table(surveys$genus)
```

#### Q6: Create new variable 'sex' by converting surveys$sex to a factor. Use addNA to add a factor level to 'sex' variable. Using the 'sex' variable, rename "NA", "F", and "M" factors to "undetermined", "female", and "male" respectively. Plot sex variable. 

```{r}
#Q6
surveys$sex <- factor(surveys$sex)
sex <- surveys$sex
sex <- addNA(sex)
levels(sex)
levels(sex)[3] <- "undetermined"
levels(sex)[2] <- "male"
levels(sex)[1] <- "female"
plot(sex)
```

#### Q7: Recreate the barplot such that "undetermined" is first (before "female").

```{r}
#Q7
sex <- factor(sex, levels = c("undetermined", "female", "male"))
plot(sex)
```

#### Q8: Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight. View the tail() of the data.frame.

```{r}
#Q8
surveys %>%
  filter(year < 1995) %>%
  select(year, sex, weight) %>%
  tail()
```

### Tidyverse Day 2

#### Q9: Create a new data frame from the surveys data that meets the following criteria:

##### Contains only the species_id column and a new column called hindfoot_cm which contains the hindfoot_length values (currently in mm) converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3.\

#### View the head() of the data.frame. 

##### Hint: Think about how the commands should be ordered to produce this data frame! 

```{r}
#Q9 

surveys <- read_csv("/Users/pvail/Documents/Courses/Comp Bio/TidyVerse/Class Repo/CM515-course-2024/modules/04_Tidyverse/diversity.csv", show_col_types = FALSE)

surveys_new <-surveys %>%
  mutate(hindfoot_cm = hindfoot_length /10) %>%
  filter(hindfoot_cm < 3) %>%
  filter(!is.na(hindfoot_length)) %>%
  select(species_id, hindfoot_cm)
surveys_new %>% head()

```

#### Q10: How many animals were caught in each plot_type surveyed?

```{r}
#Q10

surveys %>%
  #filter(plot_type != "Control")
  group_by(plot_type) %>%
  summarize("# Animals Caught" = n())
```

#### Q11: Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).

```{r}
#Q11
surveys %>%
  group_by(species_id) %>%
  filter(!is.na(hindfoot_length)) %>%
  summarize(mean_length = mean(hindfoot_length),
            min_length = min(hindfoot_length), max_length = max(hindfoot_length), n = n())

```

#### Q12: What was the heaviest animal measured in each year? Return the columns year, genus, species_id, and weight.

```{r}
#Q12

surveys %>%
  select(year, genus, species_id, weight) %>%
  filter(!is.na(weight)) %>%
  group_by(year) %>%
  filter(weight == max(weight))
  
  
```

#### Q13: Reshape the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.

```{r}
#Q13
surveys_new <- surveys %>% 
  group_by(year,plot_id) %>%
  summarize(n = n_distinct(genus)) %>%
  pivot_wider(names_from = year, values_from = n)
surveys_new
```

#### Q14: Now take that data frame and pivot_longer() it, so each row is a unique plot_id by year combination.

```{r}
#Q14
surveys_new %>%
  pivot_longer(names_to = "year", values_to = "n", cols = -plot_id)
```





